---
title: "Simulated Risk Model"
author: "James T. Bang"
date: "September 5, 2022"
output:
  pdf_document: default
  word_document: 
    reference_docx: AbstractSEA2022.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(marginaleffects); library(tidyverse); library(ggplot2)
library(stargazer); library(haven); library(interventionBias); 
library(sandwich); library(lmtest); library(parallel)
data <- read_dta('selection bias larger sample 8 8 22.dta')
data$hosp <- ifelse(data$hosp == 18, 46, data$hosp)
data$hosp <- factor(data$hosp)
data$mdc <-
  factor(
    data$mdc,
    labels = c(
      'Nervous',
      'ENT',
      'Respiratory',
      'Circulatory',
      'Digestive',
      'Musculoskeletal',
      'Skin/Subcutaneous',
      'Hepatobil./Pancreatic',
      'Endocrine/Metabolic',
      'Kidney/Urinary',
      'Male Reproductive',
      'Blood/Immune',
      'Infections/Parasitic',
      'Alcohol/Drug',
      'Injury/Poison/Toxin',
      'Missing'
    )
  )
data$age_cats <-
  factor(data$age_cats,
         labels = c('Under 45', '45-54', '55-64', '65-74', '75+'))
data$race <-
  factor(data$race, labels = c('Unspecified', 'Asian', 'Black'))
data$ethnicity <-
  factor(data$ethnicity, labels = c('Not Hispanic', 'Hispanic'))
data$gender <- factor(data$gender, labels = c('Male', 'Female'))
data$maritals <-
  factor(data$maritals, labels = c('Not Married', 'Married'))
data$hosp30d <-
  factor(data$hosp30d, labels = c('None', '30-day Prior Hosp.'))
data$medsugpt <-
  factor(data$medsugpt, labels = c('Medical', 'Surgical'))
data$icustay <- factor(data$icustay, labels = c('None', 'ICU Stay'))
data$severity <-
  factor(data$severity, labels = c('Minor', 'Moderate', 'Major'))
data$mortrisk <-
  factor(data$mortrisk, labels = c('Minor', 'Moderate', 'Major'))
data$hh <- factor(data$hh, labels = c('No HHC', 'HHC'))
data$read30_01 <-
  factor(data$read30_01, labels = c('Not Readmitted', 'Readmitted'))
n = 1000
r = 1000
outcome <- 'read30_01'
outcomeVars <-
  c(
    'hosp',
    'age_cats',
    'medsugpt',
    'icustay',
    'gender',
    'race',
    'ethnicity',
    'maritals',
    'hosp30d',
    'mortrisk',
    'severity',
    'lnlos',
    'mdc' 
  )
intervention <- 'hh'
varLabels <- c(
      read30_01 = '30-Day Readmission', 
      medsugpt = 'Surgical Patient',
      icustay = 'ICU Stay',
      gender = 'Female',
      race = 'Race',
      ethnicity = 'Ethnicity',
      maritals = 'Marital Status',
      hosp30d = 'Prior 30-Day Hospitalization',
      mortrisk = 'Mortality Risk',
      severity = 'Severity of Illness',
      mdc = 'Major Diagnostic Category',
      age_cats = 'Age',
      lnlos = 'ln(Length of Stay)',
      hh = 'Home Health')
Hmisc::label(data) <-
  as.list(varLabels[match(names(data), names(varLabels))])
outcomeVarLabels <- sapply(data, function(x)
  attr(data, 'label'))
set.seed(8675309)
```

# Data Generation and Estimation Equations 

## Data Generation

The data for observed patient outcomes of readmission, home health care (HHC) assignment, patient demographics, prior hospitalization, diagnosis, and payer type come from a single hospital in the READI study (Weiss, et al., 2019).  

Predictive models of risk attempt to assign patients into risk categories based on data known prior to discharge, and hospitals increasingly use such risk models to assign patients to post-discharge interventions like HHC. The challenge is simulating the patient's latent "risk" for the adverse outcome of readmission. We propose that this is not the same as the *ex post* *__likelihood__* of readmission. We also propose that patients with higher risk of adverse outcomes - including readmission - will be more likely to receive HHC assignments, and that the effectiveness of the clinician's decision to refer the patient to interventions like HHC at discharge masks this *ex ante* *__risk__*. Hence, there will be a bias in the using the observed likelihood to measure risk, and therefore a bias in the observed effectiveness of the intervention of assigning HHC. 

To simulate this bias, suppose that "treated" HHC patients exhibit an *ex post* readmission *prevalence*, $\pi$, but that the treatment carries a proportional *effectiveness* equal to $\eta$. Then, the overall *risk* of readmission would be: 

$$\rho = \frac{\pi}{1 - \eta}.$$

We can express the *bias* of any *ex ante* prediction based on pre-discharge observables as the true risk ($\rho$) as a proportion of the fraction of HHC patients who did *not* eventually readmit themselves. This bias, $B$ equals:

$$B = \frac{\rho}{1-\pi}.$$

Substituting (1) into (2) and simplifying, we get: 

$$B = \frac{\pi}{1-\pi} \frac{1}{1-\eta}$$



## Estimation Equations

The equations to estimate the *observed* rate of readmission and home health assignment, respectively, are:  
$$ \begin{aligned} HHC &= \beta_{10} + \beta_{11} Age + \beta_{12} Prior +  \sum_{r=1}^R \rho_{1r} Race_r + \sum_{h=1}^H \theta_{1h} Ethnicity + \sum_{p=1}^P \pi_{1p} Paytype_p \\&+ \sum_{m=1}^M \delta_{1m} Diagnosis + \alpha_{1j} + \epsilon_{1i}, \end{aligned} $$
$$ \begin{aligned} Risk &= \beta_{20} + \beta_{21} Age + \beta_{22} Prior +  \sum_{r=1}^R \rho_{2r} Race_r + \sum_{h=1}^H \theta_{2h} Ethnicity + \sum_{p=1}^P \pi_{2p} Paytype_p \\ &+ \sum_{m=1}^M \delta_{2m} Diagnosis + \alpha_{2j} + \epsilon_{2i}, \end{aligned}$$
and 
$$ \begin{aligned} Readmission &= \beta_{30} + \beta_{31} Age + \beta_{32} Prior + \beta_{33} HHC + \sum_{r=1}^R \rho_{3r} Race_r + \sum_{h=1}^H \theta_{3h} Ethnicity \\&+ \sum_{p=1}^P \pi_{3p} Paytype_p + \sum_{m=1}^M \delta_{3m} Diagnosis + \alpha_{3j} + \epsilon_{3i}, \end{aligned}$$
where $\alpha_j$ represents the fixed effect for hospital *j*. 

# Simulation Results. 

```{r, include = FALSE}
hhcSim = interventionBias::interventionSim(
  data = data,
  outcome = outcome,
  intervention = intervention,
  outcomeVars = outcomeVars
)
```

We simulated the model 1000 times using 1000 observations randomly selected from the dataset described above in each iteration. We present the replicated results for the impact of length of stay on *likelihood* of readmission and the *likelihood* of readmission in Table 1.

```{r baselineResults, results = 'asis'}
m <-
  list(
    m1 = hhcSim$results[[1]],
    m2 = hhcSim$results[[2]],
    m3 = hhcSim$results[[3]],
    m4 = hhcSim$results$Risk[1],
    m5 = hhcSim$results$Risk[2],
    m6 = hhcSim$results$Risk[3],
    m7 = hhcSim$results$Risk[4]
  )
stargazer::stargazer(
  list(m$m1, m$m2, m$m3),
  coef = hhcSim[['coefficients']][1:3],
  se = hhcSim[['se']][1:3],
  header = FALSE,
  no.space = TRUE,
  omit = c('hosp', 'mdc'),
  type = 'latex'
)
```

We present the replicated results for the impact of length of stay on *risk* of readmission of readmission in Table 2.

```{r riskResults, results = 'asis'}
stargazer::stargazer(
  list(m$m4, m$m5, m$m6, m$m7),
  coef = hhcSim[['coefficients']][4:7],
  se = hhcSim[['se']][4:7],
  header = FALSE,
  no.space = TRUE,
  omit = c('hosp', 'mdc'),
  type = 'latex'
)
```


Figure 1 presents a kernel density plot of the distributions of the two outcomes.  

```{r densityPlots, echo = FALSE}
plot.biasSim(
  hhcSim,
  'lnlos',
  title = "Figure 1: Distribution of Length of Stay Coefficient",
  xlab = "Effect of 1% Increase in Length of Stay",
  labels = c(
    "Outcome",
    "Outcome with Intervention",
    "Risk - 10% Bias",
    "Risk - 20% Bias",
    "Risk - 30% Bias",
    "Risk - 40% Bias"
  )
)
```

